Manuscript

### 2018-10-23 manuscript.Rmd
knit updated manuscript.Rmd

---
title: "Manuscript"
author: "Sally Claridge"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    code_folding: hide
    pandoc_args: [
      "+RTS", "-K2048m",
      "-RTS"
    ]
---

@@

@@

g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
# g2p_genes_filt <- c("KRAS", "EGFR", "MET", "ALK", "BRAF", "ERBB2", "ABL1", "ROS1", "KIT", "DPYD", "PDGFRA", "RET", "BRCA1", "UGT1A1", "IDH2", "ESR1", "TPMT", "BRCA2", "PDGFRB", "G6PD", "PML", "TSC1", "FLT3", "ERBB3", "CYP19A1", "AKT1", "PTCH1", "MGMT", "DNMT3A", "MYD88", "PDGFB", "TSC2", "NPM1", "JAK2")
```

@@

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
```

@@

```{r fig.height=15, fig.width=15, eval=FALSE}
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
crispr_data_signif_g2p_lineage$Hugo_Symbol <- ordered(crispr_data_signif_g2p_lineage$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_lineage_color <- as.character(crispr_data_signif_g2p_lineage$Color_Nonsilent)
names(crispr_data_signif_g2p_lineage_color) <- crispr_data_signif_g2p_lineage$Mutation_Status_Nonsilent
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_lineage, crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

@@

```{r, eval=FALSE}
ccle_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ccle_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p <- adj_signif(ccle_signif_g2p)
ccle_signif_g2p <- ccle_signif_g2p[order(ccle_signif_g2p$p),]
saveRDS(ccle_signif_g2p, "./data_munging/rds/ccle_signif_g2p_gene.rds")
```

@@

@@

```{r, eval=FALSE}
ctrp_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p <- adj_signif(ctrp_signif_g2p)
ctrp_signif_g2p <- ctrp_signif_g2p[order(ctrp_signif_g2p$p),]
saveRDS(ctrp_signif_g2p, "./data_munging/rds/ctrp_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")

@@

@@

```{r, eval=FALSE}
gdsc_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p <- adj_signif(gdsc_signif_g2p)
gdsc_signif_g2p <- gdsc_signif_g2p[order(gdsc_signif_g2p$p),]
saveRDS(gdsc_signif_g2p, "./data_munging/rds/gdsc_signif_g2p_gene.rds")
```

```{r, eval=TRUE}
gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")

@@

@@

```{r fig.height=5, fig.width=12, eval=FALSE}
crispr_data_g2p_color <- as.character(crispr_data_g2p$Color_Nonsilent)
names(crispr_data_g2p_color) <- crispr_data_g2p$Mutation_Status_Nonsilent

@@

crispr_data_g2p_plot <- ggplot(data = crispr_data_g2p, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 2) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_ptmuts_g2p_filt_color) +
  guides(color = FALSE) +
  geom_text(data = crispr_label_text, mapping = aes(x = 1.5, y = 2, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CRISPR", fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_data_g2p_plot

@@

## CCLE

@@

## CTRP

@@

## GDSC

>>>>

---
title: "Manuscript"
author: "Sally Claridge"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: true
    code_folding: hide
    pandoc_args: [
      "+RTS", "-K2048m",
      "-RTS"
    ]
---

@@

## Make per-drug boxplots
```{r}
makeDrugBoxplots <- function(drug, dataset) {
  if(dataset == "CCLE" | dataset == "ccle") {
    use_data <- filter(ccle_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ccle_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "CTRP" | dataset == "ctrp") {
    use_data <- filter(ctrp_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- ctrp_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else if(dataset == "GDSC" | dataset == "gdsc") {
    use_data <- filter(gdsc_data_g2p_genesfilt, Drug == drug)
    use_color <- as.character(use_data$Color_Nonsilent)
    names(use_color) <- use_data$Mutation_Status_Nonsilent
    sig <- gdsc_signif_g2p_ByDrug[[drug]]
    label_text <- data.frame(p.signif = sig$p.signif, p.signif.adj = sig$p.signif.adj, Hugo_Symbol = sig$Hugo_Symbol)
    label_text$Hugo_Symbol <- factor(label_text$Hugo_Symbol, levels = g2p_genes)
  }
  else { return("Error: Invalid dataset.") }

  plot <- ggplot(data = use_data, aes(x = Mutation_Status_Nonsilent, y = AUC)) +
    facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
    geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85), outlier.shape = 3, outlier.size = 0.5) +
    scale_fill_manual(values = use_color) +
    guides(color = FALSE) +
    geom_text(data = label_text, mapping = aes(x = 1.5, y = max(use_data$AUC), label = p.signif), nudge_y = 0.2) +
    theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
    labs(fill = "Mutation Status", y = "AUC", x = "Mutation Status", title = paste0(toupper(dataset), ": ", drug))
  return(plot)
}
```

@@

g2p_genes
g2p_drugs
# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("KRAS", "EGFR", "MET", "ALK", "BRAF", "ERBB2", "ABL1", "ROS1", "KIT", "DPYD", "PDGFRA", "RET", "BRCA1", "UGT1A1", "IDH2", "ESR1", "TPMT", "BRCA2", "PDGFRB", "G6PD", "PML", "TSC1", "FLT3", "ERBB3", "CYP19A1", "AKT1", "PTCH1", "MGMT", "DNMT3A", "MYD88", "PDGFB", "TSC2", "NPM1", "JAK2")
```

@@

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_drug_genes)
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_drug_genes)

# Filter drug screen datasets for G2P genes in drug-gene associations
ccle_data_g2p_genesfilt <- filter(ccle_data_ptmuts, Hugo_Symbol %in% g2p_genes)
ccle_data_g2p_genesfilt$Hugo_Symbol <- ordered(ccle_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
ctrp_data_g2p_genesfilt <- filter(ctrp_data_ptmuts, Hugo_Symbol %in% g2p_genes)
ctrp_data_g2p_genesfilt$Hugo_Symbol <- ordered(ctrp_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
gdsc_data_g2p_genesfilt <- filter(gdsc_data_ptmuts, Hugo_Symbol %in% g2p_genes)
gdsc_data_g2p_genesfilt$Hugo_Symbol <- ordered(gdsc_data_g2p_genesfilt$Hugo_Symbol, levels = g2p_genes)
```

@@

```{r fig.height=15, fig.width=15}
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_gene_filt, crispr_signif_g2p_lineage, by = c("Hugo_Symbol", "group_general_lineage_name"))
crispr_data_signif_g2p_lineage$Hugo_Symbol <- ordered(crispr_data_signif_g2p_lineage$Hugo_Symbol, levels = crispr_signif_g2p_20gene)
crispr_data_signif_g2p_lineage_color <- as.character(crispr_data_signif_g2p_lineage$Color_Nonsilent)
names(crispr_data_signif_g2p_lineage_color) <- crispr_data_signif_g2p_lineage$Mutation_Status_Nonsilent
crispr_data_signif_g2p_lineage <- merge(crispr_data_signif_g2p_lineage, crispr_data_signif_g2p_lineage %>% group_by(Hugo_Symbol) %>% summarize(Text_y = max(Score)), by = "Hugo_Symbol")

@@

```{r, eval=FALSE}
ccle_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ccle_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ccle_signif_g2p <- adj_signif(ccle_signif_g2p)
ccle_signif_g2p <- ccle_signif_g2p[order(ccle_signif_g2p$p),]
saveRDS(ccle_signif_g2p, "./data_munging/rds/ccle_signif_g2p_gene.rds")

ccle_signif_g2p_ByDrug <- lapply(unique(ccle_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = ccle_data_g2p_genesfilt, data_name = "ccle")
names(ccle_signif_g2p_ByDrug) <- unique(ccle_data_g2p_genesfilt$Drug)
saveRDS(ccle_signif_g2p_ByDrug, "./data_munging/rds/ccle_signif_g2p_ByDrug.rds")
```

@@

```{r}
ccle_signif_g2p_ByDrug <- readRDS("./data_munging/rds/ccle_signif_g2p_ByDrug.rds")
ccle_signif_g2p_ByDrug_all <- rbindlist(ccle_signif_g2p_ByDrug, use.names = TRUE)
ccle_signif_g2p_ByDrug_all <- ccle_signif_g2p_ByDrug_all[order(ccle_signif_g2p_ByDrug_all$p),]

knitr::kable(ccle_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CCLE: by-drug Wilcoxon test results for Level A point mutations, (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

```{r, eval=FALSE}
ctrp_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = ctrp_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
ctrp_signif_g2p <- adj_signif(ctrp_signif_g2p)
ctrp_signif_g2p <- ctrp_signif_g2p[order(ctrp_signif_g2p$p),]
saveRDS(ctrp_signif_g2p, "./data_munging/rds/ctrp_signif_g2p_gene.rds")

ctrp_signif_g2p_ByDrug <- lapply(unique(ctrp_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = ctrp_data_g2p_genesfilt, data_name = "CTRP")
names(ctrp_signif_g2p_ByDrug) <- unique(ctrp_data_g2p_genesfilt$Drug)
saveRDS(ctrp_signif_g2p_ByDrug, "./data_munging/rds/ctrp_signif_g2p_ByDrug.rds")
```

```{r}
ctrp_signif_g2p <- readRDS("./data_munging/rds/ctrp_signif_g2p_gene.rds")

@@

```{r}
ctrp_signif_g2p_ByDrug <- readRDS("./data_munging/rds/ctrp_signif_g2p_ByDrug.rds")
ctrp_signif_g2p_ByDrug_all <- rbindlist(ctrp_signif_g2p_ByDrug, use.names = TRUE)
ctrp_signif_g2p_ByDrug_all <- ctrp_signif_g2p_ByDrug_all[order(ctrp_signif_g2p_ByDrug_all$p),]

knitr::kable(ctrp_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "CTRP: by-drug Wilcoxon test results for Level A point mutations, (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

```{r, eval=FALSE}
gdsc_signif_g2p <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Drug_Gene", data = gdsc_data_g2p, method = "wilcox.test", p.adjust.method = "BH")
gdsc_signif_g2p <- adj_signif(gdsc_signif_g2p)
gdsc_signif_g2p <- gdsc_signif_g2p[order(gdsc_signif_g2p$p),]
saveRDS(gdsc_signif_g2p, "./data_munging/rds/gdsc_signif_g2p_gene.rds")

gdsc_signif_g2p_ByDrug <- lapply(unique(gdsc_data_g2p_genesfilt$Drug), WilcoxonByDrug, dataset = gdsc_data_g2p_genesfilt, data_name = "gdsc")
names(gdsc_signif_g2p_ByDrug) <- unique(gdsc_data_g2p_genesfilt$Drug)
saveRDS(gdsc_signif_g2p_ByDrug, "./data_munging/rds/gdsc_signif_g2p_ByDrug.rds")
```

```{r}
gdsc_signif_g2p <- readRDS("./data_munging/rds/gdsc_signif_g2p_gene.rds")

@@

```{r}
gdsc_signif_g2p_ByDrug <- readRDS("./data_munging/rds/gdsc_signif_g2p_ByDrug.rds")
gdsc_signif_g2p_ByDrug_all <- rbindlist(gdsc_signif_g2p_ByDrug, use.names = TRUE)
gdsc_signif_g2p_ByDrug_all <- gdsc_signif_g2p_ByDrug_all[order(gdsc_signif_g2p_ByDrug_all$p),]

knitr::kable(gdsc_signif_g2p_ByDrug_all[, c("Hugo_Symbol", "Drug", "p", "p.adj", "p.format", "p.signif", "p.signif.adj")], caption = "GDSC: by-drug Wilcoxon test results for Level A point mutations, (BH-adjusted p-values: * p <= 0.05, ** p <= 0.01, *** p <= 0.001, **** p <= 0.0001)") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% scroll_box(width = "900px", height = "450px")
```

@@

## Figure 1

Boxplots ordered by gene based on frequency of G2P associations.

### CRISPR

```{r fig.height=4, fig.width=20}
crispr_data_g2p$Hugo_Symbol <- factor(crispr_data_g2p$Hugo_Symbol, levels = g2p_genes)
crispr_data_g2p_color <- as.character(crispr_data_g2p$Color_Nonsilent)
names(crispr_data_g2p_color) <- crispr_data_g2p$Mutation_Status_Nonsilent

@@

crispr_data_g2p_plot <- ggplot(data = crispr_data_g2p, aes(x = Mutation_Status_Nonsilent, y = Score)) +
  facet_wrap(~ Hugo_Symbol, drop = FALSE, nrow = 1) +
  geom_hline(yintercept = 0, lty = 2, color = "darkgray") +
  geom_boxplot(mapping = aes(fill = Mutation_Status_Nonsilent), position = position_dodge(0.85)) +
  scale_fill_manual(values = crispr_data_g2p_color) +
  guides(color = FALSE) +
  geom_text(data = crispr_label_text, mapping = aes(x = 1.5, y = 2, label = p.signif), nudge_y = 0.1) +
  theme(legend.position = "top", axis.ticks.x = element_blank(), axis.text.x = element_blank(), axis.title.x = element_blank()) +
  labs(title = "CRISPR", fill = "Mutation Status", y = "CERES Score", x = "Mutation Status")
crispr_data_g2p_plot

@@

### CCLE

Plot code:
```{r, eval=FALSE}
ccle_drugs <- as.character(unique(ccle_data_g2p_genesfilt$Drug))
ccle_g2p_drugboxplots <- lapply(ccle_drugs, makeDrugBoxplots, dataset = "CCLE")
ccle_g2p_drugboxplots_paths <- paste0(ccle_drugs, "_ccle_drugboxplots.png")
pwalk(list(ccle_g2p_drugboxplots_paths, ccle_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/ccle_g2p_drugboxplots", dpi = 450, width = 20, height = 4, units = "in")
```

Data management:
```{r, eval=TRUE}
ccle_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/ccle_g2p_drugboxplots", full.names = TRUE))
names(ccle_g2p_drugboxplots) <- str_replace_all(ccle_g2p_drugboxplots, c("_ccle_drugboxplots.png" = "", "./plots_18Q3/manuscript/ccle_g2p_drugboxplots/" = ""))
bsselect(ccle_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

### CTRP

Plot code:
```{r, eval=FALSE}
ctrp_drugs <- as.character(unique(ctrp_data_g2p_genesfilt$Drug))
ctrp_g2p_drugboxplots <- lapply(ctrp_drugs, makeDrugBoxplots, dataset = "ctrp")
ctrp_g2p_drugboxplots_paths <- paste0(ctrp_drugs, "_ctrp_drugboxplots.png")
pwalk(list(ctrp_g2p_drugboxplots_paths, ctrp_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/ctrp_g2p_drugboxplots", dpi = 450, width = 20, height = 4, units = "in")
```

Data management:
```{r, eval=TRUE}
ctrp_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/ctrp_g2p_drugboxplots", full.names = TRUE))
names(ctrp_g2p_drugboxplots) <- str_replace_all(ctrp_g2p_drugboxplots, c("_ctrp_drugboxplots.png" = "", "./plots_18Q3/manuscript/ctrp_g2p_drugboxplots/" = ""))
bsselect(ctrp_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

### GDSC

Plot code:
```{r, eval=FALSE}
gdsc_drugs <- as.character(unique(gdsc_data_g2p_genesfilt$Drug))
gdsc_g2p_drugboxplots <- lapply(gdsc_drugs, makeDrugBoxplots, dataset = "gdsc")
gdsc_g2p_drugboxplots_paths <- paste0(gdsc_drugs, "_gdsc_drugboxplots.png")
pwalk(list(gdsc_g2p_drugboxplots_paths, gdsc_g2p_drugboxplots), ggsave, path = "./plots_18Q3/manuscript/gdsc_g2p_drugboxplots", dpi = 450, width = 20, height = 4, units = "in")
```

Data management:
```{r, eval=TRUE}
gdsc_g2p_drugboxplots <- paste0(list.files("./plots_18Q3/manuscript/gdsc_g2p_drugboxplots", full.names = TRUE))
names(gdsc_g2p_drugboxplots) <- str_replace_all(gdsc_g2p_drugboxplots, c("_gdsc_drugboxplots.png" = "", "./plots_18Q3/manuscript/gdsc_g2p_drugboxplots/" = ""))
bsselect(gdsc_g2p_drugboxplots, type = "img", live_search = TRUE, show_tick = TRUE, height = 300, frame_height = 275)
```

## Gene-drug associations only

### CCLE

@@

### CTRP

@@

### GDSC
